Contextual recruitment of selective attention can be updated via changes in task-relevance

### 2019-07-19 manuscript_pretty.Rmd
working - incomplete

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information [@abrahamse_grounding_2016; @brosowsky_memory-guided_2018; @crump_learning_2016; @egner_creatures_2014]. Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. How associations between attentional priorities and context cues are formed and used still remains unclear however. The current study addresses two related issues.

First, how is attentional control resolved when multiple contextual features could be used as cues? Second, it remains unclear how

Historically, theories of automaticity and attentional control have emphasized the role of practice and likened the learning process to the automaticity of skill-learning ().

However, the obligatory nature of memory-based attentional control remains unclear, in particular in situations when multiple contextual features could be used as cues for retrieval.

<!-- 
Furthermore, understanding how associations are formed and used is not well understood. Theories of automaticity and attentional control historically have emphasized the role of practice - difficult to suppress once learned.. Shiffrin and Schneider, Treisman, etc.. everyone basically.

Early CSPC work too thought that practice is important for "associations to form"

Memory-based theories on the other hand, take a more nuanced view. Practice is only important to the extent that improves memory retrieval. If retrieval cues can --  distinguish between prior episodes.

Therefore, the skill-learning view of automaticity would predict that associations take time to develop through practice, and once learned, difficult to suppress. 
-->

For example, real-world environments offer a plethora of context cues. Consider writing an article in a coffee shop. The coffee shop is a complex environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop), and associated attentional priorities. Although a laptop might cue attentional processing helpful for writing, other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory and a writer's attention would be at the mercy of the environment. On the other hand, there is some evidence that contextual cuing of attention is not entirely obligatory and can be constrained by task goals, allowing the laptop to take priority as a contextual cue over other contextual cues that are irrelevant to the task at hand.

Prior work shows that task-relevance is necessary for establishing and using context-attention associations. First, the most commonly used context cue is location [@brosowsky_context-specific_2016; @corballis_independent_2003; @crump_context-specific_2006; @crump_context-specific_2008; @crump_learning_2016; @weidler_attentional_2018; @weidler_transfer_2016]. Although in some sense, the location is irrelevant for identifying the color of a word, it seems disingenuous to consider it completely task-irrelevant. In order to identify the target, the participant must first identify and orient to its location on the screen. In this is sense, it is relevant to their ability to complete the task and furthermore, location information may receive priority during encoding [@logan_what_1998; @mayr_spatial_1996]. Similar arguments could be made about other context features that are inherent to processing the target stimulus, like font-type [@bugg_multiple_2008] and color [@vietze_context_2009].

Second, we are aware of only one study that has looked at task-relevance within the CSPC literature. Crump, Vaquero, & Milliken [-@crump_context-specific_2008] used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, @canadas_social_2013] found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for producing CSPC effects.

@@

Prior work suggests that the task-relevance of contextual information is important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects emerge when context cues are somehow made relevant to the task, but do not when made irrelevant [@canadas_social_2013; @crump_context-specific_2006; @crump_context-specific_2008]. However, it still remains unclear how task-relevance may or may not inform the use of previously learned associations. In the current study, we examined whether manipulating task-relevance would cause participants to ignore a previously learned association and adopt a new context-specific control strategy.

@@

```{r figure2b, echo=FALSE,fig.width=6.875, fig.height= 4, fig.cap='(ref:figure2b)'}
#, out.width="\\textwidth", fig.env="figure*"
#print figure 2
#source: data-analysis/analysis.R
figure2B
```

(ref:figure2b) Results from Experiment 1 showing congruency effects (incongruent - congruent) for frequency unbiased items in reaction times as a function of the task-relevant context (100% proportion congruent versus 0% proportion congruent) and phase order (100% to 0% PC versus 0% to 100% PC). Figure A shows a summary of the data containing all trials, whereas Figure B shows a summary of the data restricted to the 24 trials immediately before and 24 trials immediately after the transition. In both cases, the interaction was non-significant ($p > .05$) and Bayesian analyses showed moderate evidence in favor of the null hypothesis ($\mathrm{BF}_{\textrm{01}} > 3$).

@@

We conducted an additional, more severe test, of the list-shift hypothesis. Given that the asymmetry is interpreted as resulting from a carry-over of learned attentional modulation across phases, we might expect that any asymmetry would be exaggerated near the point of transition. We therefore repeated the above analysis but restricted it to the 24 trials immediately before the transition and 24 trials immediately after the transition.

>>>>

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information [@abrahamse_grounding_2016; @brosowsky_memory-guided_2018; @crump_learning_2016; @egner_creatures_2014]. Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. How the learning and reinstatement process works however, remains largely unclear.

The current study addresses two related issues: First, how is attentional control resolved when multiple contextual features could be used as cues? And second, what determines the expression of previously learned attentional associations?

Real-world environments can offer a plethora of context cues. For instance, consider someone writing in a coffee shop. The coffee shop is a complex environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop) associated with many experiences that afford different attentional priorities. Although a laptop might cue attentional processing helpful for writing, other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory: the mere presence of learned cues trigger attention in a ballistic manner [e.g., @schneider_controlled_1977; @shiffrin_controlled_1977]. A writer's attention in this case would be at the mercy of the environment and the most strongly associated cues. On the other hand, contextual cuing of attention may not be entirely obligatory. Our current task-goals, for instance, might constrain the expression of previously learned associations, allowing, for instance, the laptop to take priority as a contextual cue over other cues that are irrelevant to the task at hand.

Historically, perspectives of automaticity and attentional control have emphasized the role of practice and likened the learning process to the development of automaticity in skill-learning [@schneider_controlled_1977; @shiffrin_controlled_1977; @logan_toward_1988; @logan_attention_1992]. Under this view, forming an association requires many repeated experiences pairing a cue with a set of attentional priorities. Once formed however, they become habitual or hard-wired and the mere presence of the cue is suffificient to automatically adjust attention. Learned associations are therefore ballistic, difficult to suppress, and difficult to unlearn [e.g., @schneider_controlled_1977; @shiffrin_controlled_1977; @posner_attention_1975; @posner_facilitation_1975; @treisman_contextual_1960; @treisman_selective_1964; @treisman_strategies_1969]. This view is reflected in moodern work on contextual cuing where it is assumed that learned associations require practice [e.g., @lehle_fly_2008; @crump_learning_2016], form without awareness or intention, triggered automatically by the mere presence of the cue [@crump_context-specific_2006; @crump_contextual_2010; @crump_context-specific_2008; @king_priming_2012; @reuss_unconscious_2014; though, see @brosowsky_context-specific_2016].

Memory-based perspectives however, take a more nuanced view [@norman_memory_1969; @norman_memory_1976; @abrahamse_grounding_2016; @brosowsky_memory-guided_2018; @crump_learning_2016; @egner_creatures_2014]. Here, practice is only important to the extent that it improves memory retrieval. If a few prior experiences are distinct, and easily retrievable, it could support the reinstatement of a learned attentional association despite very little practice [e.g., @brosowsky_memory-guided_2018]. Conversely, if prior experiences are difficult to retrieve, then learned associations may not be expressed, even after much practice. This view offers one potential path for flexibility in the expression of learned associations. Learning an association is the process of encoding into memory experiences where attentional priorities and cues co-occur. Memory retrieval is a similarity-based, cue-driven process [e.g., @jacoby_interpreting_1978; @hintzman_minerva_1984; @hintzman_schema_1986; @hintzman_judgments_1988]. Presumably, the features of our immediate environment automatically cue memory and reinstate the retrieved attentional priorites [@logan_toward_1988; @logan_attention_1992]. However, to the extent that you could manipulate memory retrieval, you should be able manipulate the expression of learned associations.

Here, we examine whether task-relevance influences the retrieval of previously used attentional priorities. We hypothesized that the cue-driven, memory-retrieval process might be constrained to the features of our environment that are relevant to our on-going task and goals. If this is the case, then a previously learned association should be easily ignored if that feature was made irrelevant to the task because that feature would no longer contribute to memory-retrieval. Furthermore, it would suggest that the mere presence of a cue is not sufficient for automatically directing attention, but instead requires some task-integration.

Recent work lends some credence to the possibility that task-relevance is necessary for establishing and using context-attention associations. The most commonly used context cue is location [@brosowsky_context-specific_2016; @corballis_independent_2003; @crump_context-specific_2006; @crump_context-specific_2008; @crump_learning_2016; @weidler_attentional_2018; @weidler_transfer_2016], often claimed to be irrelevant to the on-going conflict task. Although in some sense, the location is irrelevant for identifying the color of a word in a Stroop task, it seems disingenuous to consider it completely task-irrelevant. In order to identify the target, the participant must first identify and orient to its location on the screen. In this is sense, it is relevant to their ability to complete the task. Further, there is some evidence that location information may receive priority during encoding [@logan_what_1998; @mayr_spatial_1996]. Similar arguments could be made about other context features that are inherent to processing the target stimulus, like font-type [@bugg_multiple_2008] and color [@vietze_context_2009].

We are aware of only one study that has looked at task-relevance within the CSPC literature. Crump, Vaquero, & Milliken [-@crump_context-specific_2008] used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, @canadas_social_2013 found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for producing CSPC effects.

@@

Prior work suggests that the task-relevance of contextual information could be important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects tend to emerge when context cues are somehow made relevant to the task, but do not when made irrelevant [@canadas_social_2013; @crump_context-specific_2006; @crump_context-specific_2008]. However, what determines the learning and expression of context-attention associations remains unclear. In the current study, we examined whether manipulating task-relevance would allow participants to ignore a previously learned association and adopt a new context-specific control strategy.

@@

```{r figure2b, echo=FALSE,fig.width=6.875, fig.height= 4, fig.cap='(ref:figure2b)'}
#, out.width="\\textwidth", fig.env="figure*"
#print figure 2
#source: data-analysis/analysis.R
#figure2B
```

(ref:figure2b) Results from Experiment 1 showing congruency effects (incongruent - congruent) for frequency unbiased items in reaction times as a function of the task-relevant context (100% proportion congruent versus 0% proportion congruent) and phase order (100% to 0% PC versus 0% to 100% PC). Figure A shows a summary of the data containing all trials, whereas Figure B shows a summary of the data restricted to the 24 trials immediately before and 24 trials immediately after the transition. In both cases, the interaction was non-significant ($p > .05$) and Bayesian analyses resulted in moderate evidence in favor of the null hypothesis ($\mathrm{BF}_{\textrm{01}} > 3$).

@@

We also conduncted an additional, more severe, test of the list-shift hypothesis. Given that the asymmetry is interpreted as resulting from a carry-over of learned attentional modulation across phases, we might expect that any asymmetry would be exaggerated near the point of transition. We therefore repeated the above analysis but restricted it to the 24 trials immediately before the transition and 24 trials immediately after the transition.

### 2019-07-19 manuscript_pretty.tex
working - incomplete

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information (Abrahamse, Braem, Notebaert, \& Verguts, 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014). Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. How associations between attentional priorities and context cues are formed and used still remains unclear however. The current study addresses two related issues.

First, how is attentional control resolved when multiple contextual features could be used as cues? Second, it remains unclear how

Historically, theories of automaticity and attentional control have emphasized the role of practice and likened the learning process to the automaticity of skill-learning ().

However, the obligatory nature of memory-based attentional control remains unclear, in particular in situations when multiple contextual features could be used as cues for retrieval.

For example, real-world environments offer a plethora of context cues. Consider writing an article in a coffee shop. The coffee shop is a complex environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop), and associated attentional priorities. Although a laptop might cue attentional processing helpful for writing, other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory and a writer's attention would be at the mercy of the environment. On the other hand, there is some evidence that contextual cuing of attention is not entirely obligatory and can be constrained by task goals, allowing the laptop to take priority as a contextual cue over other contextual cues that are irrelevant to the task at hand.

Prior work shows that task-relevance is necessary for establishing and using context-attention associations. First, the most commonly used context cue is location (Brosowsky \& Crump, 2016; Corballis \& Gratton, 2003; Crump, 2016; Crump et al., 2006; Crump, Vaquero, \& Milliken, 2008; Weidler \& Bugg, 2016; Weidler et al., 2018). Although in some sense, the location is irrelevant for identifying the color of a word, it seems disingenuous to consider it completely task-irrelevant. In order to identify the target, the participant must first identify and orient to its location on the screen. In this is sense, it is relevant to their ability to complete the task and furthermore, location information may receive priority during encoding (Logan, 1998; Mayr, 1996). Similar arguments could be made about other context features that are inherent to processing the target stimulus, like font-type (Bugg, Jacoby, \& Toth, 2008) and color (Vietze \& Wendt, 2009).

Second, we are aware of only one study that has looked at task-relevance within the CSPC literature. Crump, Vaquero, \& Milliken (2008) used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, Cañadas, Rodríguez-Bailón, Milliken, and Lupiáñez (2013){]} found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for producing CSPC effects.

@@

Prior work suggests that the task-relevance of contextual information is important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects emerge when context cues are somehow made relevant to the task, but do not when made irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, it still remains unclear how task-relevance may or may not inform the use of previously learned associations. In the current study, we examined whether manipulating task-relevance would cause participants to ignore a previously learned association and adopt a new context-specific control strategy.

@@

\begin{figure}
\centering
\includegraphics{manuscript_pretty_files/figure-latex/figure2b-1.pdf}
\caption{\label{fig:figure2b}Results from Experiment 1 showing congruency effects (incongruent - congruent) for frequency unbiased items in reaction times as a function of the task-relevant context (100\% proportion congruent versus 0\% proportion congruent) and phase order (100\% to 0\% PC versus 0\% to 100\% PC). Figure A shows a summary of the data containing all trials, whereas Figure B shows a summary of the data restricted to the 24 trials immediately before and 24 trials immediately after the transition. In both cases, the analyses resulted in no significant interaction (\(p > .05\)) and moderate evidence in favor of the null hypothesis (\(\mathrm{BF}_{\textrm{01}} > 3\)).}
\end{figure}

@@

Consistent with the original analysis, the effect of task-relevant context was significant,\(F(1, 129) = 8.92\), \(\mathit{MSE} = 3,255.69\), \(p = .003\), \(\hat{\eta}^2_p = .065\), 90\% CI \([0.01\), \(0.14]\), with moderate evidence in favor of the alternative, \(\mathrm{BF}_{\textrm{10}} = 8.11\) \([\pm 0.83\%]\). However, the two-way interaction between task-relevant context and phase order was non-significant, \(F(1, 129) = 2.56\), \(\mathit{MSE} = 3,255.69\), \(p = .112\), \(\hat{\eta}^2_p = .019\), 90\% CI \([0\), \(0.07]\), with moderate evidence in favor of the null hypothesis, \(\mathrm{BF}_{\textrm{01}} = 4.92\) \([\pm 2.95\%]\). Therefor, we did not find statistical support for a list-shift asymmetry. Additionally, the (non-significant) asymmetry we do see in this sample is in the opposite direction than would be predicted given the prior research (e.g., Abrahamse et al., 2013).

Additionally, we conducted a more severe test of the list-shift hypothesis. Given that the asymmetry is interpreted as resulting from a carry-over of learned attentional modulation across phases, we might expect that any asymmetry would be exaggerated near the point of transition. We therefore repeated the above analysis but restricted it to the 24 trials immediately before the transition and 24 trials immediately after the transition.

@@

@@

@@

@@

@@

@@

@@

@@

@@

>>>>

One explanation for CSPC effects is that the repeated application of attentional priorities in a particular context creates an associative link in episodic memory between the attentional control procedures and the contextual information (Abrahamse, Braem, Notebaert, \& Verguts, 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014). Once the associative link is established, processing the context is assumed to trigger the retrieval of previous experiences, automatically reinstating associated attentional priorities. As a result, automatic memory-based retrieval allows attentional priorities to be rapidly adjusted in a context-appropriate fashion. How the learning and reinstatement process works however, remains largely unclear.

The current study addresses two related issues: First, how is attentional control resolved when multiple contextual features could be used as cues? And second, what determines the expression of previously learned attentional associations?

Real-world environments can offer a plethora of context cues. For instance, consider someone writing in a coffee shop. The coffee shop is a complex environment with numerous context cues (e.g., pictures on the wall, chairs, coffee, laptop) associated with many experiences that afford different attentional priorities. Although a laptop might cue attentional processing helpful for writing, other features of the coffee shop might cue attentional processing related to other experiences like socializing and people-watching. Given that all of these cues are available to set attentional priorities, what determines which associative relationships will be used to guide attention? On the one hand, context-based retrieval of attentional priorities might be completely obligatory: the mere presence of learned cues trigger attention in a ballistic manner (e.g., Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). A writer's attention in this case would be at the mercy of the environment and the most strongly associated cues. On the other hand, contextual cuing of attention may not be entirely obligatory. Our current task-goals, for instance, might constrain the expression of previously learned associations, allowing, for instance, the laptop to take priority as a contextual cue over other cues that are irrelevant to the task at hand.

Historically, perspectives of automaticity and attentional control have emphasized the role of practice and likened the learning process to the development of automaticity in skill-learning (Logan, 1988, 1992; Schneider \& Shiffrin, 1977; Shiffrin \& Schneider, 1977). Under this view, forming an association requires many repeated experiences pairing a cue with a set of attentional priorities. Once formed however, they become habitual or hard-wired and the mere presence of the cue is suffificient to automatically adjust attention. Learned associations are therefore ballistic, difficult to suppress, and difficult to unlearn (e.g., Schneider \& Shiffrin, 1977; Posner \& Snyder, 1975a, 1975b; Shiffrin \& Schneider, 1977; Treisman, 1960, 1964, 1969). This view is reflected in moodern work on contextual cuing where it is assumed that learned associations require practice (e.g., Lehle \& Hübner, 2008; Crump, 2016), form without awareness or intention, triggered automatically by the mere presence of the cue (Crump et al., 2006; Crump \& Logan, 2010; Crump, Vaquero, \& Milliken, 2008; King et al., 2012; Reuss, Desender, Kiesel, \& Kunde, 2014; cf. Brosowsky \& Crump, 2016).

Memory-based perspectives however, take a more nuanced view (Abrahamse et al., 2016; Brosowsky \& Crump, 2018; Crump, 2016; Egner, 2014; Norman, 1969, 1976). Here, practice is only important to the extent that it improves memory retrieval. If a few prior experiences are distinct, and easily retrievable, it could support the reinstatement of a learned attentional association despite very little practice (e.g., Brosowsky \& Crump, 2018). Conversely, if prior experiences are difficult to retrieve, then learned associations may not be expressed, even after much practice. This view offers one potential path for flexibility in the expression of learned associations. Learning an association is the process of encoding into memory experiences where attentional priorities and cues co-occur. Memory retrieval is a similarity-based, cue-driven process (Hintzman, 1984, 1986, 1988; Jacoby, 1978). Presumably, the features of our immediate environment automatically cue memory and reinstate the retrieved attentional priorites (Logan, 1988, 1992). However, to the extent that you could manipulate memory retrieval, you should be able manipulate the expression of learned associations.

Here, we examine whether task-relevance influences the retrieval of previously used attentional priorities. We hypothesized that the cue-driven, memory-retrieval process might be constrained to the features of our environment that are relevant to our on-going task and goals. If this is the case, then a previously learned association should be easily ignored if that feature was made irrelevant to the task because that feature would no longer contribute to memory-retrieval. Furthermore, it would suggest that the mere presence of a cue is not sufficient for automatically directing attention, but instead requires some task-integration.

Recent work lends some credence to the possibility that task-relevance is necessary for establishing and using context-attention associations. The most commonly used context cue is location (Brosowsky \& Crump, 2016; Corballis \& Gratton, 2003; Crump, 2016; Crump et al., 2006, 2008; Weidler \& Bugg, 2016; Weidler et al., 2018), often claimed to be irrelevant to the on-going conflict task. Although in some sense, the location is irrelevant for identifying the color of a word in a Stroop task, it seems disingenuous to consider it completely task-irrelevant. In order to identify the target, the participant must first identify and orient to its location on the screen. In this is sense, it is relevant to their ability to complete the task. Further, there is some evidence that location information may receive priority during encoding (Logan, 1998; Mayr, 1996). Similar arguments could be made about other context features that are inherent to processing the target stimulus, like font-type (Bugg, Jacoby, \& Toth, 2008) and color (Vietze \& Wendt, 2009).

We are aware of only one study that has looked at task-relevance within the CSPC literature. Crump, Vaquero, \& Milliken (2008) used a Stroop prime-probe task where the distractor word is presented first followed by a color patch probe. Critically, they used the shape of the color patch probe, either a square or a circle, as the contextual dimension; one associated with a high frequency of conflict, the other with a low frequency. In one experiment participants were made aware of the shape-proportion contingencies but given no additional instructions. In the second, participants were given a secondary task to count the number of trials that contained a square. The first experiment failed to find evidence for context-specificity, whereas the second did. Although not testing task-relevance directly, Cañadas, Rodríguez-Bailón, Milliken, and Lupiáñez (2013) found similar effects. In this study they used images of male and female faces as context cues in a flanker task. When given instructions to think of the faces as members of the gender categories they found CSPC effects, but when given instructions to think of the faces as individuals they did not. These results suggest that making a context dimension relevant to the on-going task is important for producing CSPC effects.

@@

Prior work suggests that the task-relevance of contextual information could be important for establishing new associations between context cues and attentional priorities. Simply put, CSPC effects tend to emerge when context cues are somehow made relevant to the task, but do not when made irrelevant (Cañadas et al., 2013; Crump et al., 2006, 2008). However, it is still unclear whether there is any flexibility in the expression of learned assocations, and more importantly, what determines that flexibility. In the current study, we examined whether manipulating task-relevance would allow participants to ignore a previously learned association and adopt a new context-specific control strategy.

@@

@@

Consistent with the original analysis, the effect of task-relevant context was significant,\(F(1, 129) = 8.92\), \(\mathit{MSE} = 3,255.69\), \(p = .003\), \(\hat{\eta}^2_p = .065\), 90\% CI \([0.01\), \(0.14]\), with moderate evidence in favor of the alternative, \(\mathrm{BF}_{\textrm{10}} = 8.11\) \([\pm 0.83\%]\). However, the two-way interaction between task-relevant context and phase order was non-significant, \(F(1, 129) = 2.56\), \(\mathit{MSE} = 3,255.69\), \(p = .112\), \(\hat{\eta}^2_p = .019\), 90\% CI \([0\), \(0.07]\), with moderate evidence in favor of the null hypothesis, \(\mathrm{BF}_{\textrm{01}} = 4.92\) \([\pm 2.95\%]\). Therefore, we did not find statistical support for a list-shift asymmetry. Additionally, the (non-significant) asymmetrical effects we do see in this sample is in the opposite direction than would be predicted given the prior research (e.g., Abrahamse et al., 2013).

We also conduncted an additional, more severe, test of the list-shift hypothesis. Given that the asymmetry is interpreted as resulting from a carry-over of learned attentional modulation across phases, we might expect that any asymmetry would be exaggerated near the point of transition. We therefore repeated the above analysis but restricted it to the 24 trials immediately before the transition and 24 trials immediately after the transition.

@@

\leavevmode\hypertarget{ref-crump_contextual_2010}{}%
Crump, M. J., \& Logan, G. D. (2010). Contextual control over task-set retrieval. \emph{Attention, Perception, \& Psychophysics}, \emph{72}, 2047--2053. Retrieved from \url{http://link.springer.com/article/10.3758/BF03196681}

@@

\leavevmode\hypertarget{ref-hintzman_minerva_1984}{}%
Hintzman, D. L. (1984). MINERVA 2: A simulation model of human memory. \emph{Behavior Research Methods, Instruments, \& Computers}, \emph{16}(2), 96--101. Retrieved from \url{http://link.springer.com/article/10.3758/BF03202365}

\leavevmode\hypertarget{ref-hintzman_schema_1986}{}%
Hintzman, D. L. (1986). Schema abstraction in a multiple-trace memory model. \emph{Psychological Review}, \emph{93}, 411--428.

\leavevmode\hypertarget{ref-hintzman_judgments_1988}{}%
Hintzman, D. L. (1988). Judgments of frequency and recognition memory in a multiple-trace memory model. \emph{Psychological Review}, \emph{95}(4), 528.

@@

\leavevmode\hypertarget{ref-jacoby_interpreting_1978}{}%
Jacoby, L. L. (1978). On interpreting the effects of repetition: Solving a problem versus remembering a solution. \emph{Journal of Verbal Learning and Verbal Behavior}, \emph{17}(6), 649--667.

@@

\leavevmode\hypertarget{ref-lehle_fly_2008}{}%
Lehle, C., \& Hübner, R. (2008). On-the-fly adaptation of selectivity in the flanker task. \emph{Psychonomic Bulletin \& Review}, \emph{15}(4), 814--818. Retrieved from \url{http://link.springer.com/article/10.3758/PBR.15.4.814}

@@

\leavevmode\hypertarget{ref-logan_toward_1988}{}%
Logan, G. D. (1988). Toward an instance theory of automatization. \emph{Psychological Review}, \emph{95}, 492--527. doi:\href{https://doi.org/10.1037/0033-295X.95.4.492}{10.1037/0033-295X.95.4.492}

\leavevmode\hypertarget{ref-logan_attention_1992}{}%
Logan, G. D. (1992). Attention and preattention in theories of automaticity. \emph{The American Journal of Psychology}, 317--339.

@@

\leavevmode\hypertarget{ref-norman_memory_1969}{}%
Norman, D. A. (1969). Memory while shadowing. \emph{Quarterly Journal of Experimental Psychology}, \emph{21}(1), 85--93.

\leavevmode\hypertarget{ref-norman_memory_1976}{}%
Norman, D. A. (1976). Memory and attention: An introduction to human information processing.

\leavevmode\hypertarget{ref-posner_attention_1975}{}%
Posner, M. I., \& Snyder, C. R. R. (1975a). Attention and cognitive control. In R. L. Solso (Ed.), \emph{Information processing and cognition: The Loyola symposium} (pp. 153--175). Hillsdale, NJ: Erlbaum.

\leavevmode\hypertarget{ref-posner_facilitation_1975}{}%
Posner, M. I., \& Snyder, C. R. R. (1975b). Facilitation and inhibition in the processing of signals. In P. M. A. Rabbit \& S. Dornic (Eds.), \emph{Attention and performance V} (pp. 669--682). New York, NY: Academic Press.

@@

\leavevmode\hypertarget{ref-reuss_unconscious_2014}{}%
Reuss, H., Desender, K., Kiesel, A., \& Kunde, W. (2014). Unconscious conflicts in unconscious contexts: The role of awareness and timing in flexible conflict adaptation. \emph{Journal of Experimental Psychology: General}, \emph{143}, 1701--1718. doi:\href{https://doi.org/10.1037/a0036437}{10.1037/a0036437}

@@

\leavevmode\hypertarget{ref-schneider_controlled_1977}{}%
Schneider, W., \& Shiffrin, R. M. (1977). Controlled and automatic human information processing: I. Detection, search, and attention. \emph{Psychological Review}, \emph{84}(1), 1.

\leavevmode\hypertarget{ref-shiffrin_controlled_1977}{}%
Shiffrin, R. M., \& Schneider, W. (1977). Controlled and automatic human information processing: II. Perceptual learning, automatic attending and a general theory. \emph{Psychological Review}, \emph{84}, 127--190. doi:\href{https://doi.org/10.1037/0033-295X.84.2.127}{10.1037/0033-295X.84.2.127}

@@

\leavevmode\hypertarget{ref-treisman_contextual_1960}{}%
Treisman, A. M. (1960). Contextual cues in selective listening. \emph{Quarterly Journal of Experimental Psychology}, \emph{12}(4), 242--248.

\leavevmode\hypertarget{ref-treisman_selective_1964}{}%
Treisman, A. M. (1964). Selective attention in man. \emph{British Medical Bulletin}.

\leavevmode\hypertarget{ref-treisman_strategies_1969}{}%
Treisman, A. M. (1969). Strategies and models of selective attention. \emph{Psychological Review}, \emph{76}(3), 282.
